The error in Bayesian phylogenetic reconstruction when speciation is not instantaneous

### 2018-09-05 article.tex
Explicitly mention the incipient tree RNG seeds used

We simulate protracted birth-death trees, using the PBD package (\cite{pbd}) in the R programming language (\cite{r}).
For each combination of ${\lambda, \mu, b, n}$, we generate incipient species trees with a crown age of 15 million years.
Only trees with the desired number of good taxa are kept.

@@

\begin{figure}[!htbp]
  \includegraphics[width=0.8\textwidth]{fig_general.png}
  \caption{
    nLTT statistic distribution per biological parameter set, using the
    general data set, 
    for the subset of combinations in which $b_g = b_i$, $\mu_g = \mu_i$, 
    under the (correct) assumptions of a strict clock and Jukes-Cantor site model.
  }
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=0.8\textwidth]{fig_sampling.png}
  \caption{
    nLTT statistic distribution per biological parameter set per sampling
    regime, using the data set conditioned on sampling regime having an effect, 
    for the subset of combinations in which $b_g = b_i$, $\mu_g = \mu_i$, 
    under the (correct) assumptions of a strict clock and Jukes-Cantor site model.
  }
\end{figure}

@@

\begin{figure}[!htbp]
  \includegraphics[width=0.8\textwidth]{fig_sampling_methods.png}
  \caption{
    Sampling a species tree from an incipient species tree. At the top left,
    an incipient species tree is shown, of three different good species (the
    first and second number in the taxon label) and four different subspecies (the
    third number in the taxon tabel). The other four trees are species trees,
    that use a different sampling method to determine which sub-species is picked to
    represent a good species. These are: 'Youngest', 'Oldest', 'Shortest' and
    'Longest'. With 'Youngest' the youngest sub-species is picked to represent the
    good species. With 'Oldest' the oldest sub-species is picked to represent the
    good species. 'Shortest' is the sampling method in which the sub-species are
    picked to assure the shortest branch lengths. 'Longest' is the sampling method 
    in which the sub-species are picked to assure the longest branch lengths.
  }
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=0.8\textwidth]{fig_clock_model.png}
  \caption{
    nLTT statistic distribution per biological parameter set per clock model,
    using the general data set, 
    for the subset of combinations in which $b_g = b_i$, $\mu_g = \mu_i$, 
    under the (correct) assumption of a Jukes-Cantor site model.
  }
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=0.8\textwidth]{fig_site_model.png}
  \caption{
    nLTT statistic distribution per biological parameter set per site model,
    using the general data set, 
    for the subset of combinations in which $b_g = b_i$, $\mu_g = \mu_i$, 
    under the (correct) assumption of a strict clock model.
  }
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{p{0.05\textwidth} p{0.6\textwidth} p{0.2\textwidth}}
    \hline
                          & Description & Values \\
    \hline
    \hline
    $b_g$                 & Speciation initiation rate of a good species & derived, see \ref{table:pbd_parameters} \\
    $b_i$                 & Speciation initiation rate of an incipient species & derived, see \ref{table:pbd_parameters} \\
    $\lambda$             & Speciation completion rate & 0.1, 0.3, 1.0, $10^9$ \\
    $\mu_g$               & Extinction rate of a good species & 0.0, 0.1, 0.2 \\
    $\mu_i$               & Extinction rate of an incipient species & 0.0, 0.1, 0.2 \\
    \hline
    $n$                   & Number of good taxa & 50, 100, 200 \\
    $t_c$                 & Crown age & 15 \\
    $\sigma_c$            & Standard deviation around crown age & 0.001 \\
    $M_s$                 & Sampling method & S, L, R \\
    $M_c$                 & Clock model & S, RLN \\
    $M_t$                 & Site model & JC69, GTR \\
    $r$                   & Mutation rate & $\frac{1}{15}$ \\
    $l_a$                 & DNA alignment length & $15K$ \\
    $f_i$                 & MCMC sampling interval & 1K or more \\
    $R_i$                 & RNG seed incipient tree and randomly sampled species tree & 1, 2, ... \\
    $R_a$                 & RNG seed alignment simulation & $R_i$ \\
    $R_b$                 & RNG seed BEAST2 & $R_i$ \\
    \hline
  \end{tabular}
  \caption{
    Overview of the simulation parameters. Above the horizontal line is 
    the biological parameter set. 
    The RNG seed $R_i$ is 1 for the first simulation of the general data set, 2 for the next,
    and so on, up to and including 3480. The RNG seeds for the data set investigating the effect 
    of sampling continue from there, but only those RNG seeds are used in which sampling has an effect.
    The sampling methods are abbreviated as such: 'R' denotes random
    sampling, 'S' is 'shortest' and 'L' is 'longest'. Sampling method $M_s$ is random for the general
    data set. For the data set exploring the effect of sampling, we use 'shortest'
    and 'longest' for each value of $R_i$ (which are random seeds in which sampling has an effect).
    The clock models are abbreviated as 'S' for a strict and 'RLN' for a relaxed log-normal model.
    The site models are abbreviated as 'JC69' for Jukes-Cantor (\cite{jc69}) and 'GTR' for the generalized 
    time-reversible model (\cite{gtr}).
  }
  \label{table:parameters}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

We simulate protracted birth-death trees, using the PBD package (\cite{pbd}) in the R programming language (\cite{r}).
The first tree has a random number generator seed of 1, which is incremented by 1 for each simulated tree.
For each combination of ${\lambda, \mu, b, n}$, we generate incipient species trees with a crown age of 15 million years.
Only trees with the desired number of good taxa are kept.

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_general.png}
  \caption{
    nLTT statistic distribution per biological parameter set, using the
    general data set, 
    under the (correct) assumptions of a strict clock and Jukes-Cantor site model.
  }
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_sampling.png}
  \caption{
    nLTT statistic distribution per biological parameter set per sampling
    regime, using the data set conditioned on sampling regime having an effect, 
    under the (correct) assumptions of a strict clock and Jukes-Cantor site model.
  }
\end{figure}

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_sampling_methods.png}
  \caption{
    Sampling a species tree from an incipient species tree. At the top left,
    an incipient species tree is shown, of three different good species (the
    first and second number in the taxon label) and four different subspecies (the
    third number in the taxon tabel). The other four trees are species trees,
    that use a different sampling method to determine which sub-species is picked to
    represent a good species. These are: 'Youngest', 'Oldest', 'Shortest' and
    'Longest'. With 'Youngest' the youngest sub-species is picked to represent the
    good species. With 'Oldest' the oldest sub-species is picked to represent the
    good species. 'Shortest' is the sampling method in which the sub-species are
    picked to assure the shortest branch lengths. 'Longest' is the sampling method 
    in which the sub-species are picked to assure the longest branch lengths.
  }
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_clock_model.png}
  \caption{
    nLTT statistic distribution per biological parameter set per clock model,
    using the general data set, 
    under the (correct) assumption of a Jukes-Cantor site model.
  }
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_site_model.png}
  \caption{
    nLTT statistic distribution per biological parameter set per site model,
    using the general data set, 
    under the (correct) assumption of a strict clock model.
  }
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering 
  \begin{tabular}{p{0.05\textwidth} p{0.6\textwidth} p{0.2\textwidth}}
    \hline
                          & Description & Values \\
    \hline
    \hline
    $b_g$                 & Speciation initiation rate of a good species & derived, see \ref{table:pbd_parameters} \\
    $b_i$                 & Speciation initiation rate of an incipient species & derived, see \ref{table:pbd_parameters} \\
    $\lambda$             & Speciation completion rate & 0.1, 0.3, 1.0, $10^9$ \\
    $\mu_g$               & Extinction rate of a good species & 0.0, 0.1, 0.2 \\
    $\mu_i$               & Extinction rate of an incipient species & 0.0, 0.1, 0.2 \\
    \hline
    $n$                   & Number of good taxa & 50, 100, 200 \\
    $t_c$                 & Crown age & 15 \\
    $\sigma_c$            & Standard deviation around crown age & 0.001 \\
    $M_s$                 & Sampling method & S, L, R \\
    $M_c$                 & Clock model & S, RLN \\
    $M_t$                 & Site model & JC69, GTR \\
    $r$                   & Mutation rate & $\frac{1}{15}$ \\
    $l_a$                 & DNA alignment length & $15K$ \\
    $f_i$                 & MCMC sampling interval & 1K or more \\
    $R_i$                 & RNG seed incipient tree and randomly sampled species tree & 1, 2, etc. \\
    $R_a$                 & RNG seed alignment simulation & $R_i$ \\
    $R_b$                 & RNG seed BEAST2 & $R_i$ \\
    \hline
  \end{tabular}
  \caption{
    Overview of the simulation parameters. Above the horizontal line is 
    the biological parameter set. 
    \richel{Start new}
    The RNG seed $R_i$ is 1 for the first simulation of the general data set, 2 for the next,
    and so on.  
    The RNG seeds for the data set investigating the effect 
    of sampling continue from there, but only those RNG seeds are used in which sampling has an effect.
    \richel{End new}
    The sampling methods are abbreviated as such: 'R' denotes random
    sampling, 'S' is 'shortest' and 'L' is 'longest'. Sampling method $M_s$ is random for the general
    data set. For the data set exploring the effect of sampling, we use 'shortest'
    and 'longest' for each value of $R_i$ (which are random seeds in which sampling has an effect).
    The clock models are abbreviated as 'S' for a strict and 'RLN' for a relaxed log-normal model.
    The site models are abbreviated as 'JC69' for Jukes-Cantor (\cite{jc69}) and 'GTR' for the generalized 
    time-reversible model (\cite{gtr}).
  }
  \label{table:parameters}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
